### 2020-05-08 writeup/intro.tex
intro + reorg edits

Methods that can transfer the information from one task to help other tasks have recently found many applications in computer vision, natural language processing and reinforcement learning.
In our own experience as well as discussions with collaborators, we have often observed that multi-task learning that is applied on heterogeneous data often result in suboptimal models (or negative transfer in more technical terms).

@@

>>>>

Multi-task learning is an inductive learning mechanism to improve generalization performance using related task data.
Many state-of-the-art results in computer vision and natural language processing are obtained using multi-task learning.
On the other hand, unless the structures across task data are well-understood, applying multi-task learning on several different datasets often result in suboptimal models (or negative transfer in more technical terms).

\todo{Motivate high-dimensional data.} The technical challenge is to deal with high-dimensional data, in particular when the size of the training set is only a small constant times the feature dimension.
Prior theory is unable to explain a phase transition that occurs when comparing multi-task learning to single-task learning.
In particular in Figure \ref{fig_motivation}, we observe a shift from positive transfer to negative transfer as a parameter of task relatedness.
The theory we develop will provide a precise explanation to such a phenomenon (and more).

@@

\begin{figure}[!t]
	\includegraphics[width=0.8\textwidth]{figures/model_distance_motivation.pdf}
	\caption{Phase transition as a parameter of model distance.}
	\label{fig_motivation}
\end{figure}

### 2020-05-08 writeup/technical.tex
intro + reorg edits

\section{Main Results: Reducing Variance under Model Shift Biases}\label{sec_denoise}

@@

Concretely we will consider the following problem.
\begin{align}
	f(B; W_1, \dots, W_k) = \sum_{i=1}^k \bignorm{X B W_i - Y_i}^2. \label{eq_mtl_same_cov}
\end{align}
By fixing $W_1, W_2, \dots, W_k$, we can derive a closed form solution for $B$ as
\begin{align*}
	\hat{B}(W_1, \dots, W_k) &= (X^{\top}X)^{-1} X^{\top} \bigbrace{\sum_{i=1}^k Y_i W_i^{\top}} (Z Z^{\top})^{-1} \\
	&= \sum_{i=1}^k \bigbrace{\beta_i W_i^{\top}} (ZZ^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top} \bigbrace{\sum_{i=1}^k \varepsilon_i W_i^{\top}} (ZZ^{\top})^{-1}
\end{align*}
where we denote $Z\in\real^{r\times k}$ as the $k$ vectors $W_1, W_2, \dots, W_k$ stacked together.
%Now we switch $\hat{B}$ back into equation \eqref{eq_mtl_same_cov} to
Similar to Section \ref{sec_setup}, we consider minimizing the validation loss over $W_1, W_2, \dots, W_k$ provided with $\hat{B}$.

\paragraph{Jointly minimizing over all tasks.}
Denote by $\varepsilon(W) = \sum_{i=1}^k \varepsilon_i W_i^{\top}$.
We shall decompose the validation loss $\val(\hat{B}; W_1, \dots, W_k)$ into two parts.
The first part is the model shift bias, which is equal to
\begin{align*}
	\sum_{j=1}^k \bigbrace{\bignorm{\Sigma^{1/2}\bigbrace{\sum_{i=1}^k(\beta_i W_i^{\top}) (ZZ^{\top})^{-1} W_j - \beta_j}}^2}
\end{align*}
The second part is the variance, which is equal to
\begin{align*}
	& \sum_{j=1}^k \exarg{\varepsilon_i, \forall i}{\bigbrace{\bigbrace{\sum_{i=1}^k \varepsilon_i W_i^{\top}} (ZZ^{\top})^{-1} W_j}^2} \cdot {\bigtr{\Sigma(X^{\top}X)^{-1}}} \\
	=& \sigma^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}.
\end{align*}
Therefore we shall focus on the minimizer for the model shift bias since the variance part does not depend the weights.
Let us denote $Q\in\real^{k\times k}$ where the $(i,j)$-th entry is equal to $W_i^{\top} (ZZ^{\top})^{-1} W_j$, for any $1\le i, j\le k$.
Let $B^{\star} = [\beta_1, \beta_2, \dots, \beta_k] \in\real^{p \times k}$ denote the true model parameters.
We can now write the validation loss succinctly as follows.
\begin{align*}
	\val(\hat{B}; W_1, \dots, W_k) = \sum_{j=1}^k \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} Q - \beta_j}}^2 + \sigma^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}
\end{align*}
From the above we can solve for $Q$ optimally as \todo{this}.

\paragraph{Minimizing over the target task alone.}
If we only minimize over the the validation loss for the target task, we shall get the following.
\begin{align*}
	\val_j(\hat{B} W_j) = \bignorm{\Sigma^{1/2}\bigbrace{\sum_{i=1}^k W_i^{\top} (ZZ^{\top})^{-1}W_j \beta_i - \beta_j}}^2
	+ \sigma^2 \cdot W_j^{\top} (ZZ^{\top})^{-1} W_j \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}.
\end{align*}

From the above we can obtain three conceptual insights that are consistent with Section \ref{sec_denoise} and \ref{sec_insight}.
\begin{itemize}
	\item The de-noising effect of multi-task learning.
	\item Multi-task training vs single-task training can be either positive or negative.
	\item Transfer learning is better than the other two. And the improvement over multi-task training increases as the model distances become larger.
\end{itemize}

>>>>

\section{Main Results: A Model Shift Bias versus Variance Trade-off}\label{sec_denoise}

@@
